Storing messages with multipartite neural cliques

### 2014-07-18 Articles/article.tex
///

	To provide convenient notations, we propose to linearise messages using the following function mapping $f$:
\[
f: \left\{\begin{array}{rcl}\mathcal{A}^c &\rightarrow& \{0,1\}^{c\cdot{} \ell}\\m&\mapsto& \mu \big| \forall j, \mu[(j-1)\ell+1\dots j\times\ell] = m_j\end{array}\right.\;,
\]
where $\mu[a..b]$ is the subarray of $\mu$ from coordinate $a$ to coordinate $b$.

@@

		\begin{align*}
		&\tilde{\mu}'[(j-1)l+k] =\\ \bigg \{
  			&\begin{tabular}{cl}
  			$1$ & if $s[(j-1)\ell +k] \ge  \mbox{select}(\alpha, s[(j-1)\ell+1 \hdots j\ell])$ \\
  			$0$ & otherwise 
  		         \end{tabular}\;.
  		\end{align*}
                Thus we use an Heavyside function with a parameter depending on each cluster. $\tilde{\mu}'$ is the output of the algorithm after one iteration. The process can be repeated providing $\tilde{\mu}'$ as the new input to the network. Several stopping criterion can be used~\cite{Ala}. In the remaining of this work, we will consider the number of iterations to be fixed. Simulations show that increasing the number of iterations can result in significatively better performance.

@@

	As the probability for $i$ (and $i'$) to be active in its cluster for one message $x$ is $\frac{a}{l}$ providing messages are uniformly distributed, it follows that the probability for the edge $(i,i')$ of not being added in the network for $x$ is $1 - \left(a/l\right)^2$. Since messages are independent, the density $d$, which is the probability for an edge to be in the network, can be expressed as :

	\begin{align}
		\label{formula_density}
		d = 1 - \left( 1 - \left(\frac{a}{l}\right)^2 \right)^M 
	\end{align}		
	, recalling that $M$ is the number of messages in $\mathcal{M}$.

@@

	Since unit activations are independent (as messages are themselves independent), the probability for none of the vertices of one erased cluster, excepting the correct ones, to achieve the maximum is $\left(1 - d^	{a(c-c_e)}\right)^{l-a}$.

	Scores in clusters being also independent, the probability for none of the vertices in any erased cluster, excepting the correct ones, to achieve the maximum in their respective clusters is $\left(1 - d^	{a(c-c_e)}\right)^{c_e(l-a)}$.

	Whence error rate in retrieving messages is : 
	\begin{align}	
	P_{err} = 1 -	\left(1 - d^	{a(c-c_e)}\right)^{c_e(l-a)} 
	\end{align}

@@

	The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) l^2}{2}$ bits. Since messages are not stored in an ordered fashion, the entropy of the set of messages is $\log_2(\frac{({l \choose a}^c)^M}{M!}) \mathop{=}_{M \rightarrow +\infty} M(c \log_2{l \choose a } - \log_2(M) + \frac{1}{\log 2} + o(1)) $ (thanks to  Stirling's formula). The ratio of information effectively stored in the network is $P_{retrieve} \times  \frac{2M \left(c \log_2{l \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)l^2}$.

	Ambiguities would occur in a perfect associative memory if at least two messages were identical over non-erased letters. The probability it doesn't happen is $ (1-{l \choose a}^{c - c_e})^{M-1}$.

	$\eta_M = P_{retrieve}  \frac{2 M\left(c \log_2{l \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-(\frac{1}{{l \choose a}})^{c - c_e})^{-(M-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

@@

	Knowing the density of the network, which follows formula (\ref{formula_density}), we can compute the laws of probabilities for scores achieved by different type of units in different clusters. We can divide units in 5 groups. Subscript $g$ will indicate a unit being part of the message (it should be activated), subscript $b$ the contrary (the unit should not be activated). Subscript $a$ means the unit is already activated. Subscript $c$ corresponds to a correct cluster, $e$ to an erroneous cluster.
	Correct units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$. 
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$. We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby its law is :

@@

%	\begin{align*}
%	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
%	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
%	\end{align*}
		\begin{align*}
	&P_{retrieve} = \left [ \sum_{x = 1}^{a(c-1) + \gamma} P_{ac}(x) P(n_{bc} < x)^{l-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{x = 1}^{a(c-1) +\gamma} P_e(x) P(n_{be} < x)^{l-2a} P(n_{abe} < x)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

	with the approximation that there is few activated units that should be kept active in an erronenous cluster.

@@

		$\eta_m = P_{retrieve}  \frac{2 m\left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{l \choose a}} \right )^k \left (\frac{1}{{l \choose a}} \right )^{c-k} \right ] ^{-(m-1)}$

@@

	%If $\gamma$ is big enough,
	It follows that $P(\mbox{no other vertex activated in one cluster})= \sum_{x = 1}^{a (c - c_e)} P_c(x) P(n_v < x)^{l-a}$.%[ \sum_{x = 0}^{n - 1} P(n_v = x) \right]^{l-a }$

@@

	\begin{align}
	P_{retrieve}=  \left( \sum_{x = 1}^{a ( c- c_e)} P_c(x) P(n_v <x)^{l-a} \right )^{c_e}
	\label{psi_formula_th}
	\end{align}

@@

	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
	$\eta_{m, \psi} = P_{retrieve} m  \frac{2 \left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{l \choose a}^{c - c_e})^{-(m-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

		The entropy of the messages set can be approximated as $H(\mathcal{M}) \approx m c \log_2 {l \choose a}$. From $d = 1 - \left (1- \left(\frac{a}{l}\right)^2 \right)^m$, we can deduce $m \sim \frac{\log_2(1-d)}{\log_2(1-(\frac{a}{l})^2)} \sim - \frac{l^2 \log_2(1-d) \log 2}{a^2}$.
		We set $\log_2 {l \choose a} = k c$ (otherwise density approaches $0$ or $1$, and efficiency $0$). It follows that $\eta \sim \frac{m c \log_2 {l \choose a}}{l^2 {c \choose 2}} \sim \frac{k c^2 \log_2(1-d)\log 2}{a^2 {c \choose 2}}$.

@@

		We can bound the probability $P_{\exists e}$ for a message being "wrongly present" in the network, that is to mean adding it will not add any edge to the network albeit it is not in $\mathcal{M}$. By union-bound $P_{\exists e} \le {l \choose a}^c d^{a^2 {c \choose 2}} = 2^{c \log_ 2 {l \choose a} + a^2 {c \choose 2 }\log_2(d)} \approx 2^{c^2(k + \frac{a^2}{2} \log_2 (d))}$.

>>>>

	To provide convenient notations, we propose to linearise messages using the following function mapping $f$:
\[
f: \left\{\begin{array}{rcl}\mathcal{A}^c &\rightarrow& \{0,1\}^{c\cdot{} \ell}\\m&\mapsto& \mu \,\big| \forall j, \mu[(j-1)\ell+1\dots j\cdot \ell] = m_j\end{array}\right.\;,
\]
where $\mu[a\dots b]$ is the subarray of $\mu$ from coordinate $a$ to coordinate $b$.

@@

		\begin{align*}
		&\tilde{\mu}'[(j-1)\ell+k] =\\ \bigg \{
  			&\begin{tabular}{cl}
  			$1$ & if $s[(j-1)\ell +k] \ge  \mbox{select}(\alpha, s[(j-1)\ell+1 \hdots j\ell])$ \\
  			$0$ & otherwise 
  		         \end{tabular}\;.
  		\end{align*}
                Thus we use an Heavyside function with a parameter depending on each cluster. $\tilde{\mu}'$ is the output of the algorithm after one iteration. The process can be repeated providing $\tilde{\mu}'$ as the new input to the network. Several stopping criterion can be used~\cite{Ala}. In the remaining of this work, we will consider the number of iterations to be fixed. Simulations show that increasing the number of iterations can result in significatively better performance.

@@

	As the probability for $i$ (and $i'$) to be active in its cluster for one message $\mu$ is $a/\ell$ providing messages are uniformly distributed, it follows that the probability for the edge $(i,i')$ of not being added in the network for $\mu$ is $1 - \left(a/l\right)^2$. Since messages are independent, the density $d$, which is the probability for an edge to be in the network, can be expressed as :

	\begin{align}
		\label{formula_density}
		d = 1 - \left( 1 - \left(\frac{a}{\ell}\right)^2 \right)^M 
	\end{align}		
	, recalling that $M$ is the number of messages in $\mathcal{M}$.

@@

	Since unit activations are independent (as messages are themselves independent), the probability for none of the vertices of one erased cluster, excepting the correct ones, to achieve the maximum is $\left(1 - d^	{a(c-c_e)}\right)^{\ell-a}$.

	Scores in clusters being also independent, the probability for none of the vertices in any erased cluster, excepting the correct ones, to achieve the maximum in their respective clusters is $\left(1 - d^	{a(c-c_e)}\right)^{c_e(\ell-a)}$.

	Whence error rate in retrieving messages is : 
	\begin{align}	
	P_{err} = 1 -	\left(1 - d^	{a(c-c_e)}\right)^{c_e(\ell-a)} 
	\end{align}

@@

	The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) \ell^2}{2}$ bits. Since messages are not stored in an ordered fashion, the entropy of the set of messages is $\log_2(\frac{({\ell \choose a}^c)^M}{M!}) \mathop{=}_{M \rightarrow +\infty} M(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} + o(1)) $ (thanks to  Stirling's formula). The ratio of information effectively stored in the network is $P_{retrieve} \times  \frac{2M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2}$.

	Ambiguities would occur in a perfect associative memory if at least two messages were identical over non-erased letters. The probability it doesn't happen is $ (1-{\ell \choose a}^{c - c_e})^{M-1}$.

	$\eta_M = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times (1-(\frac{1}{{\ell \choose a}})^{c - c_e})^{-(M-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

@@

	Knowing the density of the network, which follows formula (\ref{formula_density}), we can compute the laws of probabilities for scores achieved by different type of units in different clusters. We can divide units in 5 groups with 3 characteristics. Subscript $g$ will indicate a unit being part of the message (it should be activated), subscript $b$ the contrary (the unit should not be activated). Subscript $a$ means the unit is already activated. Subscript $c$ corresponds to a correct cluster, $e$ to an erroneous cluster.
	Correct units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$. 
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$. We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby its law is :

@@

%	\begin{align*}
%	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
%	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
%	\end{align*}
		\begin{align*}
	&P_{retrieve} = \left [ \sum_{x = 1}^{a(c-1) + \gamma} P_{ac}(x) P(n_{bc} < x)^{\ell-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{x = 1}^{a(c-1) +\gamma} P_e(x) P(n_{be} < x)^{\ell-2a} P(n_{abe} < x)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

	with the approximation that there is few activated units that should be kept active in an erroneous cluster.

@@

		$\eta_m = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{\ell \choose a}} \right )^k \left (\frac{1}{{ell \choose a}} \right )^{c-k} \right ] ^{-(M-1)}$

@@

	%If $\gamma$ is big enough,
	It follows that $P(\mbox{no other vertex activated in one cluster})= \sum_{x = 1}^{a (c - c_e)} P_c(x) P(n_v < x)^{\ell-a}$.%[ \sum_{x = 0}^{n - 1} P(n_v = x) \right]^{l-a }$

@@

	\begin{align}
	P_{retrieve}=  \left( \sum_{x = 1}^{a ( c- c_e)} P_c(x) P(n_v < x)^{\ell-a} \right )^{c_e}
	\label{psi_formula_th}
	\end{align}

@@

	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
	$\eta_{m, \psi} = P_{retrieve} m  \frac{2 \left(c \log_2{\ell \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{\ell \choose a}^{c - c_e})^{-(m-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

		The entropy of the messages set can be approximated as $H(\mathcal{M}) \approx M c \log_2 {\ell \choose a}$. From $d = 1 - \left (1- \left(a/\ell\right)^2 \right)^M$, we can deduce $M \sim \frac{\log_2(1-d)}{\log_2(1-(\frac{a}{\ell})^2)} \sim - \frac{\ell^2 \log_2(1-d) \log 2}{a^2}$.
		We set $\log_2 {\ell \choose a} = k c$ (otherwise density approaches $0$ or $1$, and efficiency $0$). It follows that $\eta \sim \frac{M c \log_2 {\ell \choose a}}{\ell^2 {c \choose 2}} \sim \frac{k c^2 \log_2(1-d)\log 2}{a^2 {c \choose 2}}$.

@@

		We can bound the probability $P_{\exists e}$ for a message being "wrongly present" in the network, that is to mean adding it will not add any edge to the network albeit it is not in $\mathcal{M}$. By union-bound $P_{\exists e} \le {\ell \choose a}^c d^{a^2 {c \choose 2}} = 2^{c \log_ 2 {l \choose a} + a^2 {c \choose 2 }\log_2(d)} \approx 2^{c^2(k + \frac{a^2}{2} \log_2 (d))}$.
